# Visualizing survival curves

```{r, include =FALSE}
library(survival)
library(ggplot2)
library(gridExtra)
library(survminer)

```


Let's first look at our little survival data (to see how this works) and then we will look at the actual data.  \

The following code creates the data:

```{r little data}

simple=data.frame(matrix(c(7,8,10,3,2,3,1,1,0,1,1,0),ncol=2))
colnames(simple)=c("tenure","censored")
simple

```


To calculate the survival probabilities, we will use the survfit command.

```{r survival probabilities}

simple_km=survfit(Surv(tenure, censored==1)~1,                      data = simple)
summary(simple_km)

```
Now let's plot this survival curve.
```{r, simple survival curve}

ggsurvplot(simple_km)

```
Ok, so now let's get to the real data:

```{r simple real data}

lung_simple=survfit(Surv(time, status==2)~1,                      data = lung)
ggsurvplot(lung_simple)

```

We can also do a stratified survival curve.  To do a stratified analysis, the variables must be categorical!! This is because you are creating a curve for each level of this variable. Let's take a look at the different survival curves for sex (remember 1 is male and 2 is female). 

```{r stratified}

lung_sex=survfit(Surv(time, status==2)~sex,                      data = lung)
ggsurvplot(lung_sex)

```

You can test if there is a significant difference in the survival curves by:

```{r significance testing}

survdiff(Surv(time, status==2)~sex, data=lung)

```

